Measuring Political Freedom Around the World

### 2020-11-16 slides.Rmd
added corrections

* From Data provided by the Gapminder Foundation.

@@

* filtered to only include OECD Countries from year 2000 to 2018 by necessity

@@

1. Electoral Process

@@

1. Political Participation and Pluralism

@@

1. Discretionary Political rights.

@@

- Corruption Perception Index Ranging from 0 to 100 with 0 being the most corruption perception.

@@

- Electricity use per person in Kilowatt Hours.

@@

- Gini Coefficient. A measure of a nation's wealth inequality ranging from 0 being perfectly equal distribution of income and 1 being one person has all nation's wealth.

@@

- Percentage of population That use the Internet.

@@

- Labor Force Participation.

@@

- Murders per million people.

class: center
# Inputing Median

```{r}
set.seed(69)
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")
t2[3,2] <- median(t1[,2], na.rm = TRUE)
colnames(t2) <- c("Obs. Number", "Data Point")
knitr::kable(list(t1, t2), caption = "Illustrated Example")
```

@@

data %>% 
  group_by(polrights_fh) %>% 
  add_count(polrights_fh) %>% 
  ggplot(aes(x = polrights_fh, fill = factor(polrights_fh)), color = "#3C989E") + 
  geom_bar() + 
  geom_text(data = lab, aes(x = polrights_fh, y = n, label = n),
            color = "#3C989E", size = 8, vjust = -.10) + 
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_x_continuous(breaks = 1:7) + 
  theme_test(base_size = 30) + 
#  #  theme_xaringan() +  + 
  theme(legend.position = "none") + 
  labs(y = "Number", x = "Political Freedom", title = "Distribution of Political Freedom")
```

@@

--
* Finite mean and sd
* independence
* Linearity
* Normality
* Homoskidasticity

@@

subsets %>% 
  ggplot(aes(x = df, y = adj.r.squared)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, adj), se = FALSE, span = 0.5, color = "red") + 
  labs(title = "", x = "Parameters", y = "Adjusted R Squared") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30) 
  #  #  theme_xaringan() +  +  
```

@@

subsets %>% 
  ggplot(aes(x = df, y = AIC)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, AIC), se = FALSE, span = 0.5, color = "red") + 
  labs(title = "", x = "Parameters", y = "AIC") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30) 
  #  #  theme_xaringan() +  +  
```

@@

hist <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(.resid)) + 
  geom_histogram(bins = 100) +
  scale_x_continuous(limits = c(-4,4)) + labs(y = "Density", x = "Residual", title = "Histogram")  #  #  theme_xaringan() +  +

qq <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(sample = .resid)) + 
  stat_qq() +
  stat_qq_line() + 
scale_x_continuous(limits = c(-4,4)) + 
scale_y_continuous(limits = c(-4,4)) + labs(title = "QQ Plot of Residuals ", x = "Residual", y = "Expected") #  theme_xaringan() +  
```

@@

```{r}
fit_df5 %>% 
  broom::augment() %>% 
  mutate(yhat = predict(fit_df5),
         resid = resid(fit_df5)) %>% 
  ggplot() + 
  geom_point(aes(x = yhat, y = resid)) + 
  geom_hline(yintercept = 0, color = "red") + 
  scale_y_continuous(limits = c(-4,4)) + 
  #  theme_xaringan() +  +
  labs(y = "Residual", x = "Predicted Value")
```

@@

- We identified these outliers.
- confirmed they are not errors. 
- decided to keep them because They are indicating a substancial loss in political freedom which falls outside of the scope of the model. However, is important for our reserach Q.
***
> "A safe rule frequently suggested is to discard an outlier only if there is direct evidence that it represents an error in recording, a miscalculation, a malfunctioning of equipment, or a similar type of circumstance". -Chapter 3 p.108

@@

#### Brown Forsythe Test Dropping outliers

@@

$$
\begin{align}
t(.975; 643) = 1.963
\end{align}
$$
- Conclude Error variance is not constant

@@

$$
\begin{align}
t(.975; 639) = 1.963
\end{align}
$$
- Conclude again Error variance is not constant even when outliers are removed

>>>>

* From data provided by the Gapminder Foundation.

@@

* Filtered to only include OECD Countries from year 2000 to 2018 by necessity.

@@

### 1. Electoral Process

@@

### 2. Political Participation and Pluralism

@@

### 3. Discretionary Political Rights

@@

- Corruption Perception Index ranges from 0 to 100 with 0 being the highest perception of corruption

@@

- Electricity use per person in Kilowatt Hours

@@

- Gini Coefficient: A measure of a nation's wealth inequality ranging from 0 being perfectly equal distribution of income and 1 being one person has all nation's wealth

@@

- Percentage of population That use the Internet

@@

- Labor Force Participation

@@

- Murders per million people

---
class: center

### Illustriated example Inputing Median

```{r}
set.seed(69)
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")

colnames(t2) <- c("Obs. Number", "Data Point")

kable(t1, align = "c") %>%
  kable_styling(position = "float_left", full_width = F, html_font = "Cambria") %>% 
  row_spec(3, color = "white", background = "#3C989E")

kable(t2, align = "c") %>%
  kable_styling(full_width = FALSE, position = "left", html_font = "Cambria")
```

---
class: center

### Illustriated example Inputing Median

```{r}
set.seed(69)
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")

colnames(t2) <- c("Obs. Number", "Data Point")

kable(t1, align = "c") %>%
  kable_styling(position = "float_left", full_width = F, html_font = "Cambria") %>% 
  row_spec(3, color = "white", background = "#3C989E")

kable(t2, align = "c") %>%
  kable_styling(full_width = FALSE, position = "left", html_font = "Cambria")
```

$$Median = 264.98$$
---

class: center

### Illustriated example Inputing Median

```{r}
set.seed(69)
t1 <- cbind(1:5, round(runif(5, 200, 300), 2))
t1[3,2] <- NA
t2 <- t1
colnames(t1) <- c("Obs. Number", "Data Point")
t2[3,2] <- median(t1[,2], na.rm = TRUE)
colnames(t2) <- c("Obs. Number", "Data Point")

kable(t1, align = "c") %>%
  kable_styling(position = "float_left", full_width = F, html_font = "Cambria") %>% 
  row_spec(3, color = "white", background = "#3C989E")

kable(t2, align = "c") %>%
  kable_styling(full_width = FALSE, position = "left", html_font = "Cambria") %>% 
  row_spec(3, color = "white", background = "#3C989E")
```

@@

data %>% 
  group_by(polrights_fh) %>% 
  add_count(polrights_fh) %>% 
  ggplot(aes(x = polrights_fh, fill = factor(polrights_fh)), color = "#3C989E") + 
  geom_bar() + 
  geom_text(data = lab, aes(x = polrights_fh, y = n, label = n),
            color = "#3C989E", size = 8, vjust = -.10) + 
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_x_continuous(breaks = 1:7) + 
  theme_test(base_size = 30) + 
#    + 
  theme(legend.position = "none") + 
  labs(y = "Number", x = "Political Freedom", title = "Distribution of Political Freedom")
```

@@

--
### Finite Mean and Standard Deviation
### Independence
### Linearity
### Normality
### Homoskidasticity

@@

subsets %>% 
  ggplot(aes(x = df, y = adj.r.squared)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, adj), se = FALSE, span = 0.5, color = "red") + 
  labs(title = "", x = "Parameters", y = "Adjusted R Squared") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30) 
  #    +  
```

@@

subsets %>% 
  ggplot(aes(x = df, y = AIC)) + 
  geom_point() + 
  geom_smooth(data = r, aes(df, AIC), se = FALSE, span = 0.5, color = "red") + 
  labs(title = "", x = "Parameters", y = "AIC") + 
  scale_x_continuous(breaks = 1:9) + 
  theme_test(base_size = 30) 
  #    +  
```

@@

hist <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(.resid)) + 
  geom_histogram(bins = 100) +
  scale_x_continuous(limits = c(-4,4)) + labs(y = "Density", x = "Residual", title = "Histogram")

qq <- fit_df5 %>% 
  broom::augment() %>% 
  ggplot(aes(sample = .resid)) + 
  stat_qq() +
  stat_qq_line() + 
scale_x_continuous(limits = c(-4,4)) + 
scale_y_continuous(limits = c(-4,4)) + labs(title = "QQ Plot of Residuals ", x = "Residual", y = "Expected")   
```

@@

```{r}
fit_df5 %>% 
  broom::augment() %>% 
  mutate(yhat = predict(fit_df5),
         resid = resid(fit_df5)) %>% 
  ggplot() + 
  geom_point(aes(x = yhat, y = resid)) + 
  geom_hline(yintercept = 0, color = "red") + 
  scale_y_continuous(limits = c(-4,4)) + 
  labs(y = "Residual", x = "Predicted Value")
```

@@

- We identified these outliers.
- Confirmed they are not errors. 
- Decided to keep them because they are indicating a substantial loss in political freedom which falls outside of the scope of the model. However, is important for our research question.
***
> "A safe rule frequently suggested is to discard an outlier only if there is direct evidence that it represents an error in recording, a miscalculation, a malfunctioning of equipment, or a similar type of circumstance". -Chapter 3 p.108

@@

#### Brown-Forsythe Test Dropping outliers

@@

$$
\begin{align}
t(.975; 643) = 1.963
\end{align}
$$
- Conclude error variance is not constant

@@

$$
\begin{align}
t(.975; 639) = 1.963
\end{align}
$$
- Conclude again error variance is not constant even when outliers are removed
